The simulated market environment can be used to assess the performance of a given trade execution strategy. Using this procedure, the agent can adjust their trading strategy to acquire or sell their inventory at the most favorable price before executing the trade in reality. Section \ref{ch:strategies} provides an overview of common algorithmic trade execution strategies. Section \ref{ch:sim_results} uses or simulated market environment to evaluate the performance of the TWAP and VWAP strategies as described in Subsection \ref{ch:impact-driven}.

\section{Trade Execution Strategies} \label{ch:strategies}
Trade execution strategies can be broadly classified into impact-driven and opportunistic algorithms (\cite{labadie:hal-00590283}).

\subsection{Impact-Driven Algorithms} \label{ch:impact-driven}
Impact-driven algorithms seek to minimize the market impact of a large trade by splitting it into a series of smaller trades over time. 

Time Weighted Average Price (TWAP) algorithms split a large order of size $A$ into $n$ smaller orders of size $A/n$ uniformly over a time period $T$. Modifications of the default TWAP strategy where the size and timing of orders is slightly varied are often preferred in order to reduce predictability by third parties. When simulating TWAP performance, I use the default TWAP strategy, assuming that no third party takes action upon detecting our TWAP strategy and that the market reacts as it normally would to our trades.

Volume Weighted Average Price (VWAP) algorithms are similar to TWAP algorithms, but they divide the time period $T$ into sub-periods where the total volume of market activity differs. The size of the orders during each sub-period is proportional to its volume of market activity. The volume of market activity in each sub-period is typically predicted using historical data. VWAP has the hypothetical advantage of making larger trades when there is more market activity and smaller trades when there is less, thereby minimizing price impact. However, it runs the risk of incurring significant costs if the historical data does not reflect the trade activity at the time of execution or the volume changes significantly during trades (i.e. when a price shock occurs). The VWAP trading algorithm shares the name with the VWAP benchmark used to measure the performance of trade execution (Chapter \ref{ch:intro}).

Percentage of Volume algorithms (POV), like VWAP, generate sizes of orders based on trading volume in a given period. However, they trade a fixed percentage of the current market volume in each period based on a desired participation rate. In contrast with the TWAP and VWAP algorithms that have predetermined trading schedules, POV algorithms have a dynamically determined trading schedule.

\subsection{Cost-Driven Algorithms}
Cost-driven algorithms seek to minimize transaction costs, including market impact and market timing risks. In addition to market impact consideration taken in impact-driven algorithms above, cost-driven algorithms adjust the time horizon of trades to account for market timing risks (i.e. orders for more volatile assets are executed in a shorter time horizon). Implementation Shortfall algorithms attempt to minimize the difference between the price at which the investor decides to make the trade and the average price at which the trade is executed. Market Close algorithms are similar to Implementation Shortfall algorithms, except that the benchmark is the closing price.

\subsection{Opportunistic Algorithms}
Opportunistic algorithms take advantage of favorable market conditions. Price Inline algorithms dynamically adjust trading patterns in response to the price. Liquidity Driven algorithms take into account order book depth and availability of venues for trading. Pair Trading algorithms consist of buying one asset and selling another. If the assets are correlated the risks from one asset balance out the risks from the other. Pair Trading takes advantage of mean-reverting behavior to profit. 


\section{Trade Simulation Results} \label{ch:sim_results}
\subsection{Initial LOB}
To begin with an LOB that is representative of the market, I pick the reference price to be the time weighted average price of the asset rounded to the nearest mid-price. I pick the volume at the first 10 surrounding bid and ask prices to be the time weighted average volume at those positions, and 100 elsewhere. I then maintain and update the LOB as simulation proceeds. The starting book is shown in Table \ref{tab:starting_LOB}.

\begin{table}[htbp]
\caption{Starting LOB for Simulation. Reference Price = 516.5} \label{tab:starting_LOB}
\begin{center}
\begin{tabular}{ll|ll}
\hline \hline
\multicolumn{2}{l|}{\textbf{Bids}} & \multicolumn{2}{l}{\textbf{Asks}} \\
\hline
Price        & Volume    & Price      & Volume      \\
587.93       & 516       & 517        & 495.79      \\
1425.68      & 515       & 518        & 983.24      \\
2328.27      & 514       & 519        & 1475.59     \\
2735.44      & 513       & 520        & 2170.60     \\
2578.26      & 512       & 521        & 2456.86     \\
1338.90      & 511       & 522        & 1807.16     \\
609.82       & 510       & 523        & 817.16      \\
292.86       & 509       & 524        & 420.61      \\
250.14       & 508       & 525        & 404.11      \\
287.13       & 507       & 526        & 390.99             
\end{tabular}
\end{center}
\end{table}

\subsection{TWAP Strategy}
As described in \ref{ch:impact-driven}, the TWAP strategy consists of splitting the large trade into equally-sized smaller trades uniformly across the time period. To test the TWAP strategy, I simulate a market buy order of size 40000 over a 10 minute period. An example is shown in table \ref{tab:twap_order} where we split the trade into 8 parts. As can be seen, the total order is split into 8 orders of size 5000. The trade pushes the market price up. The best bid at the time of the first order is 517 and the best bid at the time of the last order is 528 for a VWAP of 524.820 (about 8 cents more than the starting reference price).

\begin{table}[htbp]
\begin{center}
\caption{TWAP Order} \label{tab:twap_order}
\begin{tabular}{l|l|l|l}
\hline \hline
\multicolumn{4}{c}{VWAP: 524.820}                                      \\
\hline
Time                    & Market Order Size        & Price & Volume   \\
\hline
\multirow{5}{*}{66.67}  & \multirow{5}{*}{5000} & 517   & 682.696  \\
                        &                          & 518   & 428.594  \\
                        &                          & 519   & 2566.022 \\
                        &                          & 520   & 1301.644 \\
                        &                          & 521   & 21.044   \\
\hline                   
\multirow{5}{*}{133.33} & \multirow{5}{*}{5000} & 520   & 260.565  \\
                        &                          & 521   & 1454.826 \\
                        &                          & 522   & 2239.506 \\
                        &                          & 523   & 490.201  \\
                        &                          & 524   & 554.901  \\
\hline                            
\multirow{4}{*}{200.00} & \multirow{4}{*}{5000} & 522   & 62.550   \\
                        &                          & 523   & 133.966  \\
                        &                          & 524   & 3165.501 \\
                        &                          & 525   & 1637.983 \\
\hline                        
\multirow{5}{*}{266.67} & \multirow{5}{*}{5000} & 521   & 302.563  \\
                        &                          & 522   & 411.777  \\
                        &                          & 523   & 584.960  \\
                        &                          & 524   & 3541.595 \\
                        &                          & 525   & 159.105  \\
\hline                        
\multirow{3}{*}{333.33} & \multirow{3}{*}{5000} & 524   & 684.054  \\
                        &                          & 525   & 981.617  \\
                        &                          & 526   & 3334.330 \\
\hline                        
\multirow{4}{*}{400.00} & \multirow{4}{*}{5000} & 525   & 891.352  \\
                        &                          & 526   & 1403.401 \\
                        &                          & 527   & 1557.507 \\
                        &                          & 528   & 1147.739 \\
\hline                        
\multirow{4}{*}{466.67} & \multirow{4}{*}{5000} & 526   & 631.725  \\
                        &                          & 527   & 2759.918 \\
                        &                          & 528   & 1236.733 \\
                        &                          & 529   & 371.624  \\
\hline                        
\multirow{3}{*}{533.33} & \multirow{3}{*}{5000} & 528   & 461.628  \\
                        &                          & 529   & 583.475  \\
                        &                          & 531   & 3954.897
\end{tabular}
\end{center}
\end{table}

The TWAP strategy was then tested with different numbers of order splits in the trading schedule. Each trading schedule was tested 10000 times with the same starting conditions. The results are shown in Table \ref{tab:twap_strategy}.

\begin{table}[htbp]
\begin{center}
\caption{TWAP Average Performance} \label{tab:twap_strategy}
\begin{tabular}{l|l}
\hline \hline
Number of Orders & Mean VWAP (Standard Error) \\
\hline
1                & 573.481 (0.084)            \\
2                & 566.053 (0.116)            \\
3                & 558.112 (0.130)            \\
4                & 551.222 (0.139)            \\
5                & 542.895 (0.136)            \\
6                & 536.191 (0.120)            \\
7                & 531.526 (0.096)            \\
8                & 529.286 (0.081)            \\
9                & 529.482 (0.082)            \\
10               & 528.490 (0.070)            \\
100              & 529.318 (0.061)           
\end{tabular}
\end{center}
\end{table}

As can be seen for this particular trade size, splitting the trade into smaller orders significantly lowers the market impact up to  about 8 orders. The variability of performance also generally decreases as the number of orders increases (with the exception of using just one order). The simulation results are useful for an agent looking to minimize costs by finding the optimal number of orders. Increasing the number of orders reduces market impact but has drawbacks in that it may incur more commissions from the exchange and is detected more easily. 

\subsection{VWAP Strategy}
As described in \ref{ch:impact-driven}, the VWAP strategy consists of splitting trades like the TWAP strategy but determining order sizes in proportion to the volume of market trades during different time periods. To simulate an environment with varying trading activity, we split the period of 10 minutes in half, where a certain proportion $\alpha$ of the activity occurs in the first half and $1-\alpha$ of the activity occurs in the second half. 

In order to maintain the same average volume of trades over the entire period as when the trading volume does not vary, I adjust the rate of arrival $\lambda_i$ at a given position as shown in Table \ref{tab:adjusted_rates}.

\begin{table}[htbp]
\caption{Adjusted Rates for VWAP Simulation} \label{tab:adjusted_rates}
\begin{center}
\begin{tabular}{l|l}
            & Adjusted Rate             \\
\hline
First Half  & $2 \alpha \lambda_i $    \\
Second Half & $2 (1 - \alpha) \lambda_i$
\end{tabular}
\end{center}
\end{table}

In the first half of a time period $T$, the average number of arrivals will be 
$$2 \alpha \lambda_i T / 2 = \alpha \lambda_i T$$ 
and the average number of arrivals in the second half will be 
$$2 (1-\alpha) \lambda_i T / 2 = (1-\alpha) \lambda_i T$$
for a total of $$\lambda_i T$$

Thus, on average, an $\alpha$ proportion of arrivals will occur in the first half and an $1-\alpha$ proportion occurs in the second half with the same average number of arrivals as the constant rate case. The size of the orders will be adjusted in the same way when using the VWAP strategy. It should be noted that these trading conditions are somewhat unrealistic. Although it is possible that trading volume could differ significantly in a short time interval, it is unlikely that a firm would be able to accurately predict such a drastic change and act upon it. Nevertheless, this analysis shows the possibility of testing the performance of a trading strategy in a simulated environment with varying rates of market activity. Table \ref{tab:vwap_order} shows an example of the orders executed with a VWAP strategy when $\alpha = 0.75$.

\begin{table}[htbp]
\begin{center}
\caption{VWAP Order, $\alpha = 0.75$} \label{tab:vwap_order}
\begin{tabular}{l|l|l|l}
\hline \hline
\multicolumn{4}{c}{VWAP: 522.507}                                      \\
\hline
Time                    & Market Order Size        & Price & Volume   \\
\hline
\multirow{4}{*}{66.67}  & \multirow{4}{*}{7500} & 517 & 1071.610 \\
                        &                       & 519 & 1907.321 \\
                        &                       & 520 & 4116.339 \\
                        &                       & 521 & 404.731  \\
\hline
\multirow{3}{*}{133.33} & \multirow{3}{*}{7500} & 518 & 2145.428 \\
                        &                       & 521 & 4555.106 \\
                        &                       & 522 & 799.466  \\
\hline
\multirow{6}{*}{200.00} & \multirow{6}{*}{7500} & 518 & 207.043  \\
                        &                       & 519 & 856.150  \\
                        &                       & 520 & 1513.202 \\
                        &                       & 521 & 835.556  \\
                        &                       & 522 & 1043.699 \\
                        &                       & 523 & 3044.349 \\
\hline
\multirow{5}{*}{266.67} & \multirow{5}{*}{7500} & 521 & 507.424  \\
                        &                       & 522 & 2825.699 \\
                        &                       & 523 & 1860.746 \\
                        &                       & 524 & 1134.190 \\
                        &                       & 525 & 1171.941 \\
\hline
\multirow{3}{*}{333.33} & \multirow{3}{*}{2500} & 524 & 516.300  \\
                        &                       & 525 & 1676.603 \\
                        &                       & 526 & 307.097  \\
\hline
\multirow{3}{*}{400.00} & \multirow{3}{*}{2500} & 525 & 968.908  \\
                        &                       & 527 & 759.657  \\
                        &                       & 528 & 771.434  \\
\hline
\multirow{4}{*}{466.67} & \multirow{4}{*}{2500} & 525 & 357.558  \\
                        &                       & 526 & 71.188   \\
                        &                       & 527 & 807.992  \\
                        &                       & 528 & 1263.263 \\
\hline
\multirow{3}{*}{533.33} & \multirow{3}{*}{2500} & 528 & 639.090  \\
                        &                       & 529 & 117.624  \\
                        &                       & 530 & 1743.286
\end{tabular}
\end{center}
\end{table}

Table \ref{tab:vwap_averages} shows the performance of the VWAP strategy (volume-adjusted trading) vs. the TWAP strategy (no volume-adjusted trading) for various $\alpha$ and numbers of market orders. 

It can be seen that for low numbers of orders (e.g. 4 orders), TWAP performs significantly better than VWAP, and this difference becomes more pronounced as $\alpha$ increases. This difference is likely due to the fact that the starting order book is the same in each case. Since orders in the first half of the VWAP strategy are larger, they reach farther into the order book and more units are bought at an unfavorable price. These orders have a further effect of having a larger market impact and increasing the price for the second half when the smaller orders are executed. 

The difference is mitigated when the numbers of orders are increased since the smaller orders do not reach as far into the order book. However, the TWAP strategy still largely outperforms the VWAP strategy for $\alpha = 0.6$ and $\alpha = 0.7$, likely for similar reasons as described above. At $\alpha$ and the numbers of orders are high ($\alpha = 0.8$ and 16 or 20 orders, $\alpha = 0.9$ and 12, 16, or 20 orders), the VWAP strategy outperforms the TWAP strategy. These cases are ideal for the VWAP strategy, since higher market activity in the first half means that the LOB recovers more quickly to the larger orders in the first half. The orders are also small enough so that initial orders do not reach far into the LOB. Although these conditions do not completely reflect a realistic environment, it can be seen that an ideal VWAP strategy should take into account the market activity, numbers of orders, and depth of the LOB.

\begin{table}[htbp]
\begin{center}
\caption{VWAP Averages} \label{tab:vwap_averages}
\begin{tabular}{lll}
\hline \hline
\multicolumn{1}{l|}{}                  & \multicolumn{2}{l}{Mean VWAP (Standard Error)}                                 \\ \hline
\multicolumn{1}{l|}{Numbers of Orders} & \multicolumn{1}{l|}{Volume-Adjusted Trading} & No Volume-Adjusted Trading \\ \hline
\multicolumn{3}{l}{$\alpha = 0.6$}                                                                                         \\ \hline
\multicolumn{1}{l|}{4}                 & \multicolumn{1}{l|}{547.869 (0.156)}              & 545.642 (0.144)            \\
\multicolumn{1}{l|}{8}                 & \multicolumn{1}{l|}{529.340 (0.085)}              & 528.208 (0.071)            \\
\multicolumn{1}{l|}{12}                & \multicolumn{1}{l|}{528.065 (0.068)}              & 527.514 (0.059)            \\
\multicolumn{1}{l|}{16}                & \multicolumn{1}{l|}{527.674 (0.059)}              & 527.383 (0.052)            \\
\multicolumn{1}{l|}{20}                & \multicolumn{1}{l|}{527.840 (0.057)}              & 527.366 (0.050)            \\ \hline
\multicolumn{3}{l}{$\alpha = 0.7$}                                                                                         \\ \hline
\multicolumn{1}{l|}{4}                 & \multicolumn{1}{l|}{547.437 (0.158)}              & 541.349 (0.142)            \\
\multicolumn{1}{l|}{8}                 & \multicolumn{1}{l|}{529.788 (0.093)}              & 528.257 (0.067)            \\
\multicolumn{1}{l|}{12}                & \multicolumn{1}{l|}{528.086 (0.072)}              & 527.576 (0.059)            \\
\multicolumn{1}{l|}{16}                & \multicolumn{1}{l|}{527.692 (0.063)}              & 527.492 (0.055)            \\
\multicolumn{1}{l|}{20}                & \multicolumn{1}{l|}{527.500 (0.057)}              & 527.513 (0.052)            \\ \hline
\multicolumn{3}{l}{$\alpha = 0.8$}                                                                                         \\ \hline
\multicolumn{1}{l|}{4}                 & \multicolumn{1}{l|}{548.495 (0.152)}              & 537.137 (0.133)            \\
\multicolumn{1}{l|}{8}                 & \multicolumn{1}{l|}{530.943 (0.105)}              & 529.014 (0.072)            \\
\multicolumn{1}{l|}{12}                & \multicolumn{1}{l|}{528.423 (0.076)}              & 528.485 (0.063)            \\
\multicolumn{1}{l|}{16}                & \multicolumn{1}{l|}{527.707 (0.066)}              & 528.463 (0.064)            \\
\multicolumn{1}{l|}{20}                & \multicolumn{1}{l|}{527.502 (0.059)}              & 528.692 (0.063)            \\ \hline
\multicolumn{3}{l}{$\alpha = 0.9$}                                                                                         \\ \hline
\multicolumn{1}{l|}{4}                 & \multicolumn{1}{l|}{550.013 (0.148)}              & 534.170 (0.124)            \\
\multicolumn{1}{l|}{8}                 & \multicolumn{1}{l|}{531.776 (0.110)}              & 530.629 (0.077)            \\
\multicolumn{1}{l|}{12}                & \multicolumn{1}{l|}{528.574 (0.080)}              & 530.484 (0.076)            \\
\multicolumn{1}{l|}{16}                & \multicolumn{1}{l|}{527.674 (0.066)}              & 530.641 (0.075)            \\
\multicolumn{1}{l|}{20}                & \multicolumn{1}{l|}{527.373 (0.060)}              & 530.816 (0.076)           
\end{tabular}
\end{center}
\end{table}